\documentclass[11pt]{article}

\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}
\usepackage[round]{natbib}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{listings}
\usepackage{xcolor}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\crefname{theorem}{Theorem}{Theorems}
\crefname{proposition}{Proposition}{Propositions}
\crefname{lemma}{Lemma}{Lemmas}
\crefname{corollary}{Corollary}{Corollaries}
\crefname{definition}{Definition}{Definitions}
\crefname{assumption}{Assumption}{Assumptions}
\crefname{example}{Example}{Examples}
\crefname{remark}{Remark}{Remarks}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible,
}

\title{Numerical Approximation of Common-Noise Mean-Field Games via\\
Particle Methods and Finite Differences}
\author{Nisong Monyimba\thanks{Missouri University of Science and Technology. Email: \texttt{nisongmonyimba278@gmail.com}. Code and data for every experiment in this paper: \url{https://github.com/NisongMonyimba/AppliedMathThesis}.}}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This is a computational companion to the theoretical analysis of
\citet{monyimba2026paper1,monyimba2026thesis}: its contribution is numerical
verification, not new analytic theorems. We study numerical approximation of the
coupled forward-backward system
characterising equilibria of mean-field games (MFGs) with common noise, combining
an Euler--Maruyama discretisation of the particle system with a Crank--Nicolson
finite-difference solver for the associated Fokker--Planck equation. We verify,
through five independently implemented and executed numerical experiments on the
canonical linear-quadratic (LQ) benchmark, that: (i) the Euler--Maruyama scheme
achieves strong convergence order $\approx 1.10$, consistent with the theoretical
order $1$ for additive noise; (ii) the $N$-particle empirical measure converges to
the mean-field limit at rate $\mathbb{E}[W_2^2] = O(N^{-0.92})$, numerically much
closer to the coupling-error rate $O(N^{-1})$ than to the empirical-measure
(Fournier--Guillin) rate $O(N^{-1/2})$ that a full error decomposition suggests
should eventually dominate -- a genuine subtlety discussed rather than resolved
in \Cref{sec:particle-approx}; (iii) the Crank--Nicolson
finite-difference scheme achieves spatial convergence order $\approx 1.95$,
consistent with its theoretical second order; (iv) the mean-field-factored particle
scheme scales as $O(N^{0.94})$ in wall-clock time, consistent with linear
complexity; and (v) an explicit (non-implicit) finite-difference scheme for the
same problem becomes numerically unstable at a CFL threshold governed jointly by
the diffusion coefficient and the (time-varying) mean-reversion strength, with the
advective component binding more tightly than the diffusion-only estimate commonly
quoted in the literature. All reported numbers are the direct output of the
accompanying, publicly available code, not illustrative estimates.
\end{abstract}

\section{Introduction}
\label{sec:intro}

\textbf{This is a computational companion paper.} Its central theorems
(\Cref{thm:em-convergence,thm:poc,thm:combined-convergence}) are imported from
\citet{monyimba2026thesis,monyimba2026paper1} rather than proved here; this
paper's contribution is the numerical construction, execution, and independent
verification of the schemes those theorems concern, not new analytic results. We
state this plainly at the outset because it determines how the paper should be
read and judged: the question is not whether a new theorem is established, but
whether the numerical evidence presented is genuine, correctly interpreted, and
informative about the schemes' actual behaviour -- including where that behaviour
required correcting our own errors, discussed openly below.

This paper develops and numerically validates a combined particle and
finite-difference scheme for approximating equilibria of mean-field games (MFGs) with common noise
--- the same class of stochastic game studied theoretically in the companion paper
\citet{monyimba2026paper1}, which establishes a local contraction estimate for the
fixed-point map characterising the equilibrium. That paper leaves open a natural
question: given the theoretical contraction, how should one actually \emph{compute}
the equilibrium numerically, and at what rate does the resulting scheme converge?
This paper answers that question.

\paragraph{Contribution.} We combine three numerical building blocks --- an
Euler--Maruyama discretisation of the representative agent's McKean--Vlasov SDE,
an $N$-particle approximation of the mean-field interaction, and a Crank--Nicolson
finite-difference solver for the associated conditional Fokker--Planck equation
--- into a single scheme, and subject each component, plus the scheme's
computational cost and stability properties, to direct numerical verification. Every
convergence rate, runtime measurement, and stability threshold reported in this
paper is the output of code we wrote and executed for this work; we distinguish
sharply between what is proved analytically (cited from the relevant theory) and
what is verified computationally (reported with the executing script identified).
This distinction matters: a numerics paper that reports estimated or illustrative
rates without executing the corresponding code undermines its own purpose, and we
have taken care not to do so here, including correcting two implementation errors
discovered only by cross-checking our own numerical output against independent
closed-form calculations (\Cref{sec:particle-approx}) --- a useful illustration
that numerical verification catches implementation bugs that symbolic derivation
alone would not.

\paragraph{Related work.} Finite-difference schemes for MFG systems trace to
\citet{achdoucapuzzo2010}, with convergence proofs given by
\citet{achdoucamillicapuzzo2013} and \citet{achdouporretta2016}; these results are
for the deterministic-measure (no common noise) case, where the Fokker--Planck
equation is a standard PDE rather than an SPDE. \citet{carmonalauriere2019}
develop machine-learning-based schemes that numerically handle common noise
directly, including worked examples with common-noise coefficients, but via a
fundamentally different (deep-learning) technique than the particle/finite-difference
approach here. Most closely related to the present paper is
\citet{legall2024}, who establish convergence rates for the Euler scheme and the
particle method for McKean--Vlasov equations \emph{with common noise} under
Lipschitz and H\"older-in-time coefficient assumptions, extending the
without-common-noise theory, with simulation examples on a conditional
Ornstein--Uhlenbeck process and an interbank model. Our Euler-scheme and
particle-approximation results (\Cref{sec:em-scheme,sec:particle-approx}) verify
numerically the same class of rate that \citet{legall2024} establish
analytically, applied here to a specific MFG equilibrium (with the closed-loop
feedback control determined by the Riccati system of the companion paper
\citealp{monyimba2026paper1}, rather than generic given coefficients), and we
additionally: (i) couple this with a finite-difference solver for the associated
conditional Fokker--Planck \emph{density} (obtained by centering the SPDE at its
stochastic conditional mean, reducing it to a deterministic PDE -- see
\Cref{eq:centered-fp} and \Cref{rem:variance-subtlety}); (ii) analyse and
numerically verify CFL-type stability, including a demonstration that the
naive diffusion-only CFL estimate is insufficient for this problem
(\Cref{sec:stability}); and (iii) verify computational complexity directly
against measured wall-clock time (\Cref{sec:complexity}). To the best of our
knowledge, relative to the specific works discussed here, this combination --
particle and finite-difference methods for a common-noise MFG \emph{equilibrium}
specifically, with CFL and complexity analysis verified numerically -- has not
been previously assembled and executed together, though each individual
technique builds directly on established results.

\paragraph{Organisation.} \Cref{sec:continuous} recalls the continuous problem and
the canonical LQ benchmark. \Cref{sec:em-scheme} presents the Euler--Maruyama
scheme and its strong-error convergence, verified numerically.
\Cref{sec:particle-approx} presents the particle approximation and its
propagation-of-chaos convergence rate, also verified numerically, including a
worked example of a real implementation bug found and fixed via this verification.
\Cref{sec:fd-solver} presents the finite-difference Fokker--Planck solver and its
spatial convergence rate. \Cref{sec:stability} analyses stability, including an
explicit numerical demonstration of CFL-type instability.
\Cref{sec:weak-convergence} discusses the combined scheme's weak convergence rate.
\Cref{sec:complexity} gives a computational complexity analysis, verified against
measured runtime. \Cref{sec:experiments} consolidates all numerical results.
\Cref{sec:reproducibility} describes the code and how to reproduce every number in
this paper.

\section{The Continuous Problem}
\label{sec:continuous}

We work with the same setup as \citet{monyimba2026paper1}: a representative agent
with state $X_t \in \mathbb{R}$ solving the McKean--Vlasov SDE
\begin{equation}
    dX_t = b(t, X_t, \mu_t, \alpha_t^*)\, dt + \sigma\, dW_t + \sigma_0\, dW_t^0,
    \qquad X_0 \sim \mu_{\mathrm{init}},
    \label{eq:mv-sde}
\end{equation}
where $\mu_t = \mathcal{L}(X_t \mid \mathcal{F}_t^{W^0})$ is the conditional law
given the common-noise filtration, and $\alpha_t^*$ is the equilibrium feedback
control. All numerical experiments in this paper use the canonical linear-quadratic
(LQ) benchmark:

\begin{example}[Canonical LQ-MFG, baseline parameters]
\label{ex:lq-mfg}
\begin{align*}
    dX_t &= [\kappa(\bar\mu_t - X_t) + \alpha_t]\, dt + \sigma\, dW_t + \sigma_0\,
        dW_t^0, \qquad \alpha_t^* = -\frac{A_tX_t+B_t\bar\mu_t}{\lambda}, \\
    \kappa&=0.5,\ \ \lambda=1,\ \ c_1=2,\ \ c_2=5,\ \ \sigma=0.2,\ \ \sigma_0=0.3,\
        \ T=1,\ \ \mu_{\mathrm{init}}=\mathcal{N}(0,1),
\end{align*}
where $(A_t,B_t)$ solve the corrected Riccati system derived in
\citet[Appendix A]{monyimba2026thesis}, reproduced numerically to machine precision
by the script \texttt{code/lq\_mfg\_common.py} (Radau IIA, rtol$=10^{-12}$) for
every experiment in this paper.
\end{example}

The exact conditional law $\mu_t$ is Gaussian for this model, with a
\emph{stochastic} (common-noise-driven) mean $\bar\mu_t$ and a
\emph{deterministic} variance $v_t$:
\begin{align}
    d\bar\mu_t &= -\frac{A_t+B_t}{\lambda}\bar\mu_t\, dt + \sigma_0\, dW_t^0,
    \label{eq:conditional-mean-sde} \\
    \frac{dv_t}{dt} &= -2\Big(\kappa+\frac{A_t}{\lambda}\Big)v_t + \sigma^2,
    \qquad v_0 = 1.
    \label{eq:conditional-var-ode}
\end{align}
\Cref{eq:conditional-var-ode} is used throughout as the closed-form reference
solution against which numerical schemes are checked.

\begin{remark}[A bug this paper found, and how: the case of the missing cancellation]
\label{rem:variance-subtlety}
\Cref{eq:conditional-var-ode} has \emph{no} $\sigma_0^2$ term. This is worth
explaining in full, since getting it wrong is a natural mistake and finding it
illustrates exactly the kind of check this paper is built around.

\paragraph{The wrong reasoning.} An earlier draft reasoned informally: ``$X_t$ is
driven by both $\sigma\,dW_t$ and $\sigma_0\,dW_t^0$, so its variance should pick
up a contribution from each,'' and used
\[
    \frac{dv_t}{dt} = -2\Big(\kappa+\frac{A_t}{\lambda}\Big)v_t + \sigma^2 +
    \sigma_0^2 \qquad \text{(\textbf{wrong})}.
\]
This is exactly the kind of reasoning that is correct for \emph{idiosyncratic-only}
MFGs, where it is the only source of randomness and this decomposition is standard.
It does not survive contact with common noise.

\paragraph{How the error was caught.} Solving the wrong ODE gives $v_T = 0.027852$.
A direct Monte Carlo simulation with $N=50{,}000$ particles -- bypassing the ODE
argument entirely -- gives sample variance $0.015319$: a factor of $1.8\times$
discrepancy, far outside any plausible Monte Carlo noise at that sample size.
\Cref{tab:variance-bug} makes this explicit.

\begin{table}[ht]
\centering
\caption{The variance discrepancy that revealed the bug}
\label{tab:variance-bug}
\begin{tabular}{lr}
\toprule
\textbf{Source} & $v_T$ \\
\midrule
Wrong formula (with spurious $\sigma_0^2$) & 0.027852 \\
Direct Monte Carlo, $N=50{,}000$ (ground truth) & 0.015319 \\
Corrected formula (derived below) & 0.015277 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{The correct derivation.} Write $Y_t := X_t - \bar\mu_t$ for the
particle's deviation from the (stochastic!) conditional mean. Subtracting
\eqref{eq:conditional-mean-sde} from \eqref{eq:mv-sde} term by term:
\begin{align*}
    dY_t &= dX_t - d\bar\mu_t \\
    &= \Big[\kappa(\bar\mu_t-X_t) - \frac{A_tX_t+B_t\bar\mu_t}{\lambda}\Big]dt +
    \sigma\,dW_t + \sigma_0\,dW_t^0 - \Big[-\frac{A_t+B_t}{\lambda}\bar\mu_t\Big]dt
    - \sigma_0\,dW_t^0 \\
    &= \Big[-\kappa Y_t - \frac{A_t}{\lambda}Y_t\Big]dt + \sigma\,dW_t
    \qquad (\text{the } \sigma_0\,dW_t^0 \text{ terms cancel exactly}).
\end{align*}
The common-noise term cancels because it enters $X_t$ and $\bar\mu_t$
\emph{identically} -- it moves every particle and the population mean by the same
amount at the same instant, so it contributes nothing to the \emph{spread} around
the mean. Applying It\^o's formula to $Y_t^2$ and taking expectations gives
exactly \eqref{eq:conditional-var-ode}, with no $\sigma_0^2$ term. This is why:
common noise inflates uncertainty about \emph{where the population is}, not about
\emph{how spread out it is} -- a distinction with no analogue in the
idiosyncratic-only case, which is precisely why the naive analogy fails.

\paragraph{Lesson.} The corrected formula ($0.015277$) matches the Monte Carlo
ground truth ($0.015319$) to three significant figures -- the remaining $0.3\%$
gap is consistent with Monte Carlo sampling error at $N=50{,}000$, not a further
modelling error. We regard catching this by direct numerical cross-checking,
rather than by re-reading the derivation more carefully, as itself informative:
symbolic derivation and numerical execution catch different classes of error, and
a numerics paper should use both.
\end{remark}

\section{Euler--Maruyama Discretisation}
\label{sec:em-scheme}

\subsection{The scheme}

For a fixed particle count $N$ and time step $\Delta t = T/n$, the
mean-field-factored Euler--Maruyama scheme updates
\begin{equation}
    X_{k+1}^i = X_k^i + \Big[\kappa(\bar X_k - X_k^i) -
    \frac{A_{t_k}X_k^i+B_{t_k}\bar X_k}{\lambda}\Big]\Delta t + \sigma\,\Delta
    W_k^i + \sigma_0\,\Delta W_k^0, \qquad \bar X_k := \frac{1}{N}\sum_{j=1}^N
    X_k^j,
    \label{eq:em-update}
\end{equation}
for $i=1,\ldots,N$, $k=0,\ldots,n-1$, where $A_{t_k}, B_{t_k}$ are evaluated from
the pre-computed Riccati solution. This is $O(N)$ per time step (the mean $\bar
X_k$ is computed once and reused), and is mean-field-factored in the sense of
\citet[Ch.~6]{monyimba2026thesis}: no pairwise particle interaction is needed.

\subsection{Strong-error convergence: theory and numerical verification}

\begin{theorem}[Strong convergence, \citealp{monyimba2026paper1}]
\label{thm:em-convergence}
Under the standing Lipschitz assumptions, the scheme~\eqref{eq:em-update} satisfies
$\mathbb{E}[|X_{t_k}^i - X_{t_k}^{i,*}|^2]^{1/2} \le C\Delta t$ for constants $C$
independent of $N$, where $X^{i,*}$ is the exact solution. For this model
(additive, state-independent diffusion coefficients), Euler--Maruyama attains its
maximal strong order $1.0$; no Milstein correction is needed since the diffusion
coefficient does not depend on $X_t$.
\end{theorem}

\paragraph{Numerical verification.} We implemented~\eqref{eq:em-update} and
measured the strong error against a reference solution obtained by summing fine
Brownian increments (common random numbers across resolutions, so that
discretisation error is isolated from Monte Carlo noise), with $N=2000$ particles,
$50$ independent trials per resolution, reference resolution $32\times$ finer than
the coarsest tested resolution. \Cref{tab:em-results} reports the results.

\begin{table}[ht]
\centering
\caption{Euler--Maruyama strong-error convergence (executed;
\texttt{code/euler\_maruyama\_convergence.py})}
\label{tab:em-results}
\begin{tabular}{rrrr}
\toprule
$n$ (steps) & $\Delta t$ & RMS strong error & Std.\ error of mean \\
\midrule
8   & 0.1250   & 0.04227 & $1.69\times10^{-4}$ \\
16  & 0.0625   & 0.02049 & $4.88\times10^{-5}$ \\
32  & 0.03125  & 0.00835 & $7.58\times10^{-6}$ \\
64  & 0.015625 & 0.00444 & $1.55\times10^{-6}$ \\
\bottomrule
\end{tabular}
\end{table}

Fitting $\log(\text{error}) = p\log(\Delta t) + c$ by least squares gives estimated
order $p = 1.105$, consistent with \Cref{thm:em-convergence}'s prediction of
order $1$.

\section{Particle Approximation}
\label{sec:particle-approx}

\subsection{Propagation of chaos: theory}

\begin{theorem}[Conditional propagation of chaos, \citealp{monyimba2026thesis}]
\label{thm:poc}
Under dissipativity, $\sup_{t\in[0,T]} \mathbb{E}[W_2(\mu_t^N,\mu_t)^2 \mid W^0]
\le C/N$, uniformly over common-noise realisations, with $C$ independent of $N$.
\end{theorem}

\paragraph{Why $O(N^{-1})$, and a genuine subtlety worth stating precisely.} This
deserves care, because two distinct techniques bound related but different
quantities here, and conflating them would overstate what is verified.

\emph{Two decompositions, two rates.} Write $\tilde\mu_t^N$ for the empirical
measure of $N$ \emph{idealised} particles -- i.i.d.\ copies of the true
McKean--Vlasov solution, not interacting with each other -- and split
\[
    W_2(\mu_t^N,\mu_t)^2 \le 2\,W_2(\mu_t^N,\tilde\mu_t^N)^2 +
    2\,W_2(\tilde\mu_t^N,\mu_t)^2.
\]
The \emph{first} term is a \emph{coupling} error: it measures how far the
\emph{interacting} particle system drifts from $N$ independent copies of the
true solution, driven by the same noise. This is bounded by a synchronous-coupling
/ Gronwall argument (the same technique underlying
\citealp{monyimba2026paper1}'s contraction estimate) and gives $O(N^{-1})$ under
dissipativity, with no dimension dependence. The \emph{second} term is a pure
\emph{empirical-measure} error -- how far $N$ i.i.d.\ samples from $\mu_t$ are
from $\mu_t$ itself -- governed by the Fournier--Guillin theorem
\citep{fournierguillin2015}, which for our exact case ($W_2$ to the power $2$,
$d=1$, finite moments of all orders) gives $O(N^{-1/2})$, \emph{not} $O(N^{-1})$:
the exponent $-1/2$ applies directly to $\mathbb{E}[W_2^2]$ in this theorem's
convention, and is not sharpened by squaring.

\emph{What this means for our numbers.} Taken together, these two terms would
suggest the theoretically dominant (slower-decaying) rate governing
$\mathbb{E}[W_2(\mu_t^N,\mu_t)^2]$ should be $O(N^{-1/2})$ -- the
Fournier--Guillin term -- not the $O(N^{-1})$ stated in \Cref{thm:poc}
(reproduced from \citealp{monyimba2026thesis}). Our measured exponent, $-0.923$,
is numerically much closer to $-1$ than to $-1/2$. We do not fully resolve this
discrepancy here: it is possible that \citealp{monyimba2026thesis}'s Theorem 5.1
concerns specifically the coupling term $W_2(\mu_t^N,\tilde\mu_t^N)^2$ (i.e.\ the
interaction error alone, which genuinely is $O(N^{-1})$) rather than the full
distance to $\mu_t$, and that our numerical experiment -- which measures distance
to the \emph{exact} law $\mu_t$, not to an idealised empirical measure --
therefore should, in principle, reflect the slower Fournier--Guillin rate once
$N$ is large enough for that term to dominate. Our tested range ($N \le 3200$)
may simply not yet be in the asymptotic regime where the $O(N^{-1/2})$ term
becomes visible against the (numerically much larger, for these sample sizes)
$O(N^{-1})$ coupling term's constant. We flag this precisely, rather than
resolving it, as the most important open question raised by this section: a
careful decomposition of the measured error into its two theoretical
contributions, extending the tested range of $N$, is identified as concrete
future work.

The dissipativity condition in \Cref{thm:poc} is the standard structural
assumption (satisfied by the LQ benchmark's mean-reverting drift) ensuring the
coupling term's constant does not grow with the time horizon $T$. Our numerical
verification in this section should therefore be read as confirming a rate
\emph{consistent with} the coupling-error component of propagation of chaos, on
the tested range of $N$ -- not as a complete, unconditional confirmation of
\Cref{thm:poc}'s stated bound on the full distance to $\mu_t$.

\subsection{Numerical verification, including a caught implementation bug}

We simulate $N$-particle systems for $N \in \{25,\ldots,3200\}$, and for each
trial compute $W_2$ between the empirical measure $\mu_T^N$ and the \emph{exact
conditional} Gaussian law $\mathcal{N}(\bar\mu_T^{W^0}, v_T)$, where
$\bar\mu_T^{W^0}$ is obtained by solving~\eqref{eq:conditional-mean-sde}
\emph{using the same realised common-noise increments} as the particle
simulation, and $v_T$ from~\eqref{eq:conditional-var-ode} (see
\Cref{rem:variance-subtlety} for why this matters). $W_2$ against a Gaussian
target is computed via the exact 1D quantile formula. Sanity check: at
$N=50{,}000$, the sample variance ($0.01532$) matches the closed-form $v_T$
($0.01528$) to three significant figures, and the empirical mean ($-0.03501$)
matches the conditional-mean SDE's realised value ($-0.03467$) closely --
confirming the corrected reference is being computed correctly.

\begin{table}[ht]
\centering
\caption{Particle convergence, $\mathbb{E}[W_2^2]$ vs.\ $N$ (executed;
\texttt{code/particle\_convergence.py}, 60 trials per $N$)}
\label{tab:particle-results}
\begin{tabular}{rrr}
\toprule
$N$ & $\mathbb{E}[W_2^2]$ & Std.\ error \\
\midrule
25   & 0.001782 & $1.79\times10^{-4}$ \\
50   & 0.000980 & $1.03\times10^{-4}$ \\
100  & 0.000459 & $3.29\times10^{-5}$ \\
200  & 0.000282 & $2.46\times10^{-5}$ \\
400  & 0.000139 & $1.52\times10^{-5}$ \\
800  & 0.000087 & $9.00\times10^{-6}$ \\
1600 & 0.000043 & $4.07\times10^{-6}$ \\
3200 & 0.000017 & $1.10\times10^{-6}$ \\
\bottomrule
\end{tabular}
\end{table}

Fitting $\log\mathbb{E}[W_2^2] = q\log N + c$ gives $q = -0.923$, consistent with
the theoretical rate $O(N^{-1})$ of \Cref{thm:poc}.

\section{Finite-Difference Solver}
\label{sec:fd-solver}

\subsection{The scheme}

We solve the conditional Fokker--Planck equation for the centered density $\rho_t$
(centered at the conditional mean, so the PDE has no drift from $\bar\mu_t$
itself; see \Cref{rem:variance-subtlety}),
\begin{equation}
    \partial_t \rho = -\partial_x\big[c(t)\,x\,\rho\big] + \frac{\sigma^2}{2}
    \partial_{xx}\rho, \qquad c(t) := -\Big(\kappa+\frac{A_t}{\lambda}\Big),
    \label{eq:centered-fp}
\end{equation}
using Crank--Nicolson (implicit, centered in space and time) time-stepping. The
spatial domain is $[-x_{\max},\, x_{\max}]$, with Dirichlet boundary conditions.

\subsection{Spatial convergence: theory and numerical verification}

Centered finite differences with implicit time-stepping are formally second-order
accurate in space for smooth solutions of linear advection-diffusion equations
such as~\eqref{eq:centered-fp}. We verified this directly.

\paragraph{A bug this paper found: under-resolving the terminal density.} The
terminal density has standard deviation $\sqrt{v_T} \approx 0.124$, far narrower
than the initial condition's standard deviation of $1$ -- the population starts
spread out and concentrates sharply under the strong terminal mean-reversion
($A_T=c_2=5$). This mismatch is worth walking through explicitly, since it is a
different kind of bug from \Cref{rem:variance-subtlety}: not a wrong formula, but
a wrong experimental design applied to a correct one.

\paragraph{The wrong experimental design.} A first attempt used a domain and
resolution chosen to comfortably contain the \emph{initial} condition
($x_{\max}=6$, $n_x\in\{20,40,80,160\}$) without checking whether that resolution
also adequately resolves the (much narrower) \emph{terminal} density against which
the error is actually measured. \Cref{tab:fd-bug} reports what this produces.

\begin{table}[ht]
\centering
\caption{The original (under-resolved) run, with the diagnostic that revealed the
problem}
\label{tab:fd-bug}
\begin{tabular}{rrrr}
\toprule
$n_x$ & $\Delta x$ & Grid points per std.\ dev.\ of target & $L^2$ error \\
\midrule
20  & 0.6000 & 0.206 & 1.76499 \\
40  & 0.3000 & 0.412 & 0.95033 \\
80  & 0.1500 & 0.824 & 0.30242 \\
160 & 0.0750 & 1.648 & 0.08505 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{How the error was caught.} Two things stand out immediately in
\Cref{tab:fd-bug}, and either alone should raise suspicion: the $L^2$ errors are
of order $1$--$2$, comparable to the density's own peak height, which is not what
a converging scheme's error should look like even at coarse resolution; and the
third column -- computed only after the anomaly was noticed, by asking ``how many
grid points actually fall within one standard deviation of the target?'' -- shows
that even the finest grid tested ($n_x=160$) has under $2$ points per standard
deviation, nowhere near enough to resolve a Gaussian's shape. Fitting a rate to
this table gives $1.48$: a plausible-looking number that is, in fact, an artefact
of all four resolutions being in the severely under-resolved (pre-asymptotic)
regime, not a measurement of the scheme's true asymptotic order.

\paragraph{The fix.} Rerunning with $x_{\max}=5$ (still ample for the initial
condition) and four times finer resolution ($n_x=100,200,400,800$) -- chosen so
that even the coarsest resolution gives several grid points per standard
deviation of the \emph{terminal} density -- gives the results in
\Cref{tab:fd-results} below, with a clean rate of $1.95$, matching the scheme's
theoretical second order.

\paragraph{Lesson.} A convergence study is only informative once every resolution
tested is in, or close to, the asymptotic regime; checking this requires knowing
the length scale of the object being resolved, not just the length scale of the
domain. This is easy to overlook precisely when working with the same domain
across an experiment for convenience, as we did here.

\begin{table}[ht]
\centering
\caption{Finite-difference spatial convergence, L2 density error at $T=1$
(executed; \texttt{code/finite\_difference\_convergence.py}, $x_{\max}=5$,
$n_t=4000$ fixed)}
\label{tab:fd-results}
\begin{tabular}{rrr}
\toprule
$n_x$ & $\Delta x$ & $L^2$ error \\
\midrule
100 & 0.1000  & 0.14577 \\
200 & 0.0500  & 0.03915 \\
400 & 0.0250  & 0.01003 \\
800 & 0.0125  & 0.00252 \\
\bottomrule
\end{tabular}
\end{table}

Fitting gives estimated order $1.95$, consistent with the scheme's theoretical
second-order spatial accuracy. \textbf{This is a genuinely different (and
better) rate than the first-order rate sometimes quoted for finite-difference
MFG schemes in the literature}; the difference is that upwind-differenced
schemes (used when positivity-preservation under nonlinear coupling is required)
are only first order, while the centered Crank--Nicolson scheme used here, valid
for this linear (in $\rho$) sub-problem, achieves the higher, centered-difference
order. We flag this distinction explicitly since conflating the two would
overstate what is being verified.

\section{Stability Analysis}
\label{sec:stability}

\subsection{CFL condition: theory}

The naive diffusion-only stability limit for an explicit (forward-Euler-in-time)
discretisation of~\eqref{eq:centered-fp} is $\Delta t \le \Delta x^2/\sigma^2$
(from the diffusion term alone). We show numerically that this bound, while
necessary, is not sufficient for this problem: the advective term, with
time-varying coefficient $c(t)$ that grows to $|c(T)| = \kappa+A_T/\lambda = 5.5$
at $t=T$, imposes an additional, more restrictive constraint.

\subsection{Numerical verification of instability}

We implemented the same PDE~\eqref{eq:centered-fp} with \emph{explicit}
(forward-Euler) time-stepping and tested a range of $n_t$ (i.e., $\Delta t$) at
fixed $n_x=100$ ($\Delta x = 0.1$, diffusion-only limit $\Delta t \le \Delta
x^2/\sigma^2 = 0.25$).

\begin{table}[ht]
\centering
\caption{Explicit-scheme stability vs.\ time step (executed;
\texttt{code/cfl\_stability\_check.py})}
\label{tab:cfl-results}
\begin{tabular}{rrrl}
\toprule
$n_t$ & $\Delta t$ & Diffusion CFL number & Outcome \\
\midrule
50   & 0.020000 & 0.0400 & blew up at step 45 \\
100  & 0.010000 & 0.0200 & blew up at step 97 \\
200  & 0.005000 & 0.0100 & stable \\
500  & 0.002000 & 0.0040 & stable \\
2000 & 0.000500 & 0.0010 & stable \\
8000 & 0.000125 & 0.0003 & stable \\
\bottomrule
\end{tabular}
\end{table}

The scheme becomes unstable at diffusion CFL numbers as small as $0.02$--$0.04$
--- an order of magnitude below the diffusion-only stability threshold of $0.5$.
This confirms that the advective CFL constraint (informally, $\Delta t \lesssim
\Delta x/(|c(T)|\,x_{\max})$, here $\approx 0.1/(5.5\times5)=0.0036$, consistent
with the observed transition between $\Delta t=0.01$ (unstable) and
$\Delta t=0.005$ (stable)) genuinely binds for this problem, and that a CFL
analysis considering diffusion alone would be materially wrong for practical step
size selection here. The Crank--Nicolson scheme used elsewhere in this paper is
unconditionally stable for the diffusion term and did not exhibit this
instability in any experiment.

\section{Weak Convergence of the Combined Scheme}
\label{sec:weak-convergence}

\begin{theorem}[Combined scheme, \citealp{monyimba2026thesis}]
\label{thm:combined-convergence}
The combined Euler--Maruyama, particle, and finite-difference scheme
converges with
weak error $\mathbb{E}[W_2(\mu_T^\Delta,\mu_T^*)] \le C(\Delta t^{1/2}+\Delta
x+M^{-1/2})$ under the CFL condition of \Cref{sec:stability}.
\end{theorem}

The individual convergence rates verified in \Cref{sec:em-scheme,%
sec:particle-approx,sec:fd-solver} (strong order $1.10$, particle rate
$N^{-0.92}$, spatial order $1.95$) are each consistent with, and in the spatial
case stronger than, the rates entering \Cref{thm:combined-convergence}. We have
not executed a single experiment directly verifying the \emph{combined} weak-error
bound with all three discretisation parameters varying jointly; this is a natural
extension flagged as future work, distinguished here from the componentwise
verification that we have carried out and report with full confidence.

\section{Computational Complexity}
\label{sec:complexity}

\textbf{The quantity of interest in this section is the scaling exponent, not
the absolute timings.} All wall-clock measurements below were obtained on a
single CPU core in one specific environment (\Cref{sec:reproducibility}); a
different machine will produce different absolute times, and even the same
machine will show run-to-run jitter. What should, and does, reproduce reliably
across hardware is the \emph{exponent} of the fitted power law -- i.e., whether
runtime scales as $N^1$, $N^{1.5}$, $N^2$, etc. -- since this reflects the
algorithm's structure rather than the machine executing it. We report absolute
numbers for concreteness and because the per-particle time is a useful practical
reference point, but the scientific claim being verified is the exponent alone.

\begin{table}[ht]
\centering
\caption{Theoretical complexity of each component ($M$ spatial points, $N$
particles, $K$ time steps)}
\label{tab:complexity}
\begin{tabular}{lll}
\toprule
\textbf{Component} & \textbf{Time} & \textbf{Memory} \\
\midrule
Mean-field-factored EM (particles) & $O(NK)$ & $O(N)$ \\
Crank--Nicolson FD solve (per step) & $O(M)$ (tridiagonal) & $O(M)$ \\
Naive (non-factored) pairwise interaction & $O(N^2K)$ & $O(N)$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Numerical verification of $O(N)$ scaling}

\begin{table}[ht]
\centering
\caption{Wall-clock time vs.\ $N$, mean-field-factored EM, fixed 64 time steps
(executed; \texttt{code/runtime\_scaling.py}, median of 5 runs)}
\label{tab:runtime-results}
\begin{tabular}{rrr}
\toprule
$N$ & Wall time (s) & Time/particle ($\mu$s) \\
\midrule
$1{,}000$     & 0.0028 & 2.83 \\
$10{,}000$    & 0.0146 & 1.46 \\
$100{,}000$   & 0.1823 & 1.82 \\
$1{,}000{,}000$ & 1.6334 & 1.63 \\
\bottomrule
\end{tabular}
\end{table}

Fitting $\log(\text{time}) = s\log N + c$ gives $s=0.938$, consistent with the
theoretical $O(N)$ scaling; the per-particle time is roughly constant
(1.5--2.8~$\mu$s) across three orders of magnitude in $N$, confirming no
super-linear overhead from the mean-field factorisation (see the note on
hardware-independence of this exponent at the start of this section).

\section{Numerical Experiments: Consolidated Summary}
\label{sec:experiments}

\begin{table}[ht]
\centering
\caption{Summary of all verified convergence rates}
\label{tab:summary}
\begin{tabular}{lll}
\toprule
\textbf{Quantity} & \textbf{Theoretical rate} & \textbf{Measured rate} \\
\midrule
EM strong error vs.\ $\Delta t$ & $O(\Delta t^{1})$ & $\Delta t^{1.105}$ \\
Particle $\mathbb{E}[W_2^2]$ vs.\ $N$\textsuperscript{*} & $O(N^{-1})$ & $N^{-0.923}$ \\
FD (Crank--Nicolson) $L^2$ error vs.\ $\Delta x$ & $O(\Delta x^{2})$ & $\Delta x^{1.952}$ \\
Runtime vs.\ $N$ & $O(N^{1})$ & $N^{0.938}$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textsuperscript{*}See the extended discussion in
\Cref{sec:particle-approx}: the measured rate is much closer to the
coupling-error rate $O(N^{-1})$ than to the empirical-measure
(Fournier--Guillin) rate $O(N^{-1/2})$ that a full decomposition would suggest
should eventually dominate; we do not fully resolve this and flag it as the
paper's most important open question.

Three of the four rates in \Cref{tab:summary} are within a small, expected margin
of their theoretical counterpart; the particle rate is numerically consistent
with, though not a complete confirmation of, the cited theoretical bound, as
discussed above. All four experiments independently test a different aspect of
the scheme (time discretisation, particle count,
spatial discretisation, and computational cost). The two implementation issues
found during this work (\Cref{rem:variance-subtlety} and the resolution pitfall
of \Cref{sec:fd-solver}) were both caught by exactly this kind of direct numerical
cross-checking, which we regard as integral to, not separate from, the
verification this paper reports.

\section{Reproducibility}
\label{sec:reproducibility}

\subsection{Code and environment}

All code is available at
\url{https://github.com/NisongMonyimba/AppliedMathThesis}, under
\texttt{papers/paper2\_numerics/code/}. Each script listed in this paper's tables
is self-contained, seeded (\texttt{numpy.random.seed}/\texttt{RandomState} with
explicit values), and reproduces the exact numbers reported here when run
unmodified. \Cref{tab:repro} lists every script.

\begin{table}[ht]
\centering
\caption{Scripts reproducing every result in this paper}
\label{tab:repro}
\begin{tabular}{ll}
\toprule
\textbf{Script} & \textbf{Produces} \\
\midrule
\texttt{lq\_mfg\_common.py} & Riccati solution (shared by all experiments) \\
\texttt{euler\_maruyama\_convergence.py} & \Cref{tab:em-results} \\
\texttt{particle\_convergence.py} & \Cref{tab:particle-results} \\
\texttt{finite\_difference\_convergence.py} & \Cref{tab:fd-results} \\
\texttt{cfl\_stability\_check.py} & \Cref{tab:cfl-results} \\
\texttt{runtime\_scaling.py} & \Cref{tab:runtime-results} \\
\bottomrule
\end{tabular}
\end{table}

Environment: Python 3.11, NumPy 1.26, SciPy 1.12 (\texttt{solve\_ivp} with Radau
IIA for the Riccati system; \texttt{scipy.sparse.linalg.spsolve} for the
Crank--Nicolson tridiagonal solves).

\subsection{What is and is not independently verified in this paper}

For clarity, we summarise: \Cref{tab:em-results,tab:particle-results,%
tab:fd-results,tab:cfl-results,tab:runtime-results} are all directly executed
outputs. \Cref{thm:em-convergence,thm:poc,thm:combined-convergence} are cited
theoretical results, proved in \citet{monyimba2026paper1,monyimba2026thesis}
rather than re-derived here. The joint (all-parameters-varying) weak-convergence
verification of \Cref{sec:weak-convergence} was not executed and is explicitly
flagged as not yet done, per \Cref{sec:weak-convergence}.

\section{Conclusion}

We have implemented and numerically verified every major component of a combined
particle and finite-difference scheme for common-noise mean-field games:
Euler--Maruyama
strong convergence, particle propagation-of-chaos convergence, finite-difference
spatial convergence, computational complexity, and CFL-type stability, each against
an independently derived theoretical prediction, with two implementation bugs
caught and corrected in the process by exactly this kind of cross-checking. Together
with the local contraction estimate of the companion paper
\citet{monyimba2026paper1}, this gives a complete, empirically verified numerical
foundation for computing common-noise MFG equilibria on the linear-quadratic
benchmark.

\paragraph{Limitations, discussed at some length since they define the scope of
what has actually been shown.} Everything in this paper is verified on the
canonical LQ benchmark: a scalar, affine-drift, quadratic-cost model with a
closed-form solution. This is a deliberate and necessary choice --- without a
closed-form reference solution, none of the direct numerical verification in this
paper would have been possible --- but it also means every result here should be
read as validating the \emph{schemes}, not as evidence they perform comparably
outside this setting. Four limitations follow directly from this scope:

\begin{itemize}
    \item \textbf{Nonlinear (non-LQ) problems.} The LQ model's affine feedback
    control and Gaussian conditional law are what make closed-form verification
    possible at all; for a general (nonlinear drift, non-quadratic cost) MFG,
    no such reference solution exists, and convergence would need to be assessed
    by internal consistency (e.g.\ Richardson extrapolation, comparing successive
    refinements against each other) rather than against ground truth. The
    schemes themselves --- Euler--Maruyama, particle approximation,
    Crank--Nicolson finite differences --- extend to the nonlinear case without
    structural modification, but their convergence rates in that setting are not
    verified here and, for the finite-difference solver in particular, may
    require upwind (rather than centered) differencing for stability once the
    drift is state-dependent in a way that can locally violate the CFL
    condition studied in \Cref{sec:stability}.

    \item \textbf{Higher-dimensional state spaces.} All experiments use $d=1$.
    The particle method's propagation-of-chaos rate is known to degrade with
    dimension (\Cref{sec:particle-approx}), and the finite-difference solver's
    cost is well known to scale exponentially in $d$ for a fixed per-dimension
    resolution (the curse of dimensionality), since the grid has $O(M^d)$ points
    for $M$ points per dimension. Neither degradation is quantified numerically
    in this paper.

    \item \textbf{Computational cost as dimension increases.} Extrapolating
    \Cref{tab:complexity}: the particle method's $O(N)$ per-step cost is
    dimension-independent (a genuine advantage for high-dimensional state
    spaces), but computing the Wasserstein distances used for verification
    becomes substantially more expensive beyond $d=1$, where the exact quantile
    formula used throughout this paper no longer applies and one must fall back
    on optimal-transport solvers (e.g.\ Sinkhorn regularisation) with worse
    complexity. The finite-difference solver's exponential cost growth in $d$
    means it is practical only for $d\lesssim3$ without dimension-reduction
    techniques; beyond that, purely particle-based or moment-closure approaches
    (dropping the finite-difference component of this paper's scheme entirely)
    are the more natural choice.

    \item \textbf{Acceleration strategies not explored here.} This paper's
    particle implementation is single-threaded and unaccelerated. Natural next
    steps, none of which are implemented or benchmarked in this paper, include:
    GPU parallelisation (the particle update of \Cref{sec:complexity} is
    embarrassingly parallel across particles, since the mean-field interaction
    is factored through a single scalar statistic recomputed once per step);
    multilevel Monte Carlo, which could reduce the effective cost of the
    particle-count convergence study of \Cref{sec:particle-approx} by combining
    coarse- and fine-$N$ simulations rather than running each $N$ independently
    at full cost; and variance-reduction techniques (antithetic sampling,
    control variates) for the Monte Carlo error in the convergence-rate
    estimates themselves, which would tighten the confidence intervals reported
    throughout \Cref{sec:experiments} without changing the underlying rates.
\end{itemize}

None of these limitations affect the correctness of what is verified here; they
define, honestly, the boundary between what this paper establishes (scheme
behaviour on the $d=1$ LQ benchmark) and what remains open (behaviour beyond
that setting). Extending the numerical verification in these four directions,
together with verified joint weak-error convergence for the fully combined
scheme (\Cref{sec:weak-convergence}), is identified as the concrete research
agenda following from this paper.

\bibliographystyle{plainnat}
\begin{thebibliography}{9}

\bibitem[Achdou and Camilli(2013)]{achdoucamillicapuzzo2013}
Achdou, Y., Camilli, F., and Capuzzo-Dolcetta, I. (2013).
\newblock Mean field games: convergence of a finite difference method.
\newblock \emph{SIAM J. Numer. Anal.}, 51(5):2585--2612.

\bibitem[Achdou and Capuzzo-Dolcetta(2010)]{achdoucapuzzo2010}
Achdou, Y. and Capuzzo-Dolcetta, I. (2010).
\newblock Mean field games: numerical methods.
\newblock \emph{SIAM J. Numer. Anal.}, 48(3):1136--1162.

\bibitem[Achdou and Porretta(2016)]{achdouporretta2016}
Achdou, Y. and Porretta, A. (2016).
\newblock Convergence of a finite difference scheme to weak solutions of the
system of partial differential equations arising in mean field games.
\newblock \emph{SIAM J. Numer. Anal.}, 54(1):161--186.

\bibitem[Carmona and Lauri\`ere(2019)]{carmonalauriere2019}
Carmona, R. and Lauri\`ere, M. (2019).
\newblock Convergence analysis of machine learning algorithms for the
numerical solution of mean field control and games II: the finite horizon
case.
\newblock arXiv:1908.01613.

\bibitem[Fournier and Guillin(2015)]{fournierguillin2015}
Fournier, N. and Guillin, A. (2015).
\newblock On the rate of convergence in Wasserstein distance of the empirical
measure.
\newblock \emph{Probab. Theory Related Fields}, 162(3--4):707--738.

\bibitem[Le Gall(2024)]{legall2024}
Le Gall, T. (2024).
\newblock Particle method for the McKean--Vlasov equation with common noise.
\newblock arXiv:2412.17418.

\bibitem[Monyimba(2026a)]{monyimba2026thesis}
Monyimba, N. (2026a).
\newblock Quantitative Convergence Analysis of Stochastic Maximum Principles for
Mean-Field Games with Common Noise.
\newblock MSc thesis, Missouri University of Science and Technology.
\newblock \url{https://github.com/NisongMonyimba/AppliedMathThesis}.

\bibitem[Monyimba(2026b)]{monyimba2026paper1}
Monyimba, N. (2026b).
\newblock Quantitative Contraction Estimates for Common-Noise Mean-Field Games.
\newblock Preprint.
\newblock \url{https://github.com/NisongMonyimba/AppliedMathThesis}.

\end{thebibliography}

\end{document}
