\begin{abstract}
\noindent
This thesis develops a mathematically rigorous and computationally implementable
framework for mean-field games with common noise, integrating three research streams
--- SPDE theory, multi-timescale reinforcement learning, and quantitative finance ---
into a single coherent pipeline. The central claim is that market inefficiency can be
productively understood as distributional disequilibrium under shared shocks: a
measurable distance in probability space between the observed distribution of
institutional investor positions and the game-theoretic equilibrium.

The work provides six principal contributions.

\begin{enumerate}
    \item \textbf{Quantitative well-posedness} (Theorem 3.1). Under standard
    regularity assumptions, the mean-field game with common noise admits a unique
    equilibrium, and the fixed-point map is a contraction in a weighted Wasserstein-2
    metric. The stability constant is explicit in the model parameters
    ($C = 2L_b^2 + 2\|\sigma_0\|_F^2 + C_{\mathrm{reg}}$), directly informing numerical
    step-size selection and reinforcement-learning rate design. The contraction is
    established in expectation under synchronous coupling of the common noise.

    \item \textbf{Numerical convergence} (Theorem 4.1). A combined Euler--Maruyama /
    particle / SPDE discretisation converges with weak error
    $O(\Delta t^{1/2} + \Delta x)$, and a CFL condition coupling parabolic, stochastic
    transport, and explicit feedback constraints is derived and validated on the
    linear-quadratic benchmark.

    \item \textbf{Three-timescale reinforcement-learning convergence} (Theorem 5.1).
    A multi-timescale actor-critic algorithm converges almost surely along
    subsequences (in the sense $\liminf W_2 = 0$) to the equilibrium measure flow,
    in continuous state-action spaces under compact Lipschitz function
    approximators. An exponential
    convergence rate under entropy regularisation and a log-Sobolev inequality is also
    established (Theorem 5.2).

    \item \textbf{SPDE-grounded hierarchical architecture} (Proposition 6.1). The
    master-equation decomposition into idiosyncratic, mean-field, and common-noise
    operators canonically motivates a three-tier architecture (SAC-PPO-TD3), which
    significantly outperforms monolithic baselines in sample efficiency and stability,
    with a control experiment isolating the SPDE-guided timescale contribution.

    \item \textbf{Alpha signals from the equilibrium} (Proposition 8.1). Four trading
    signals are derived from the linear-quadratic MFG equilibrium. Three are adaptations
    of prior work (optimal spread, mean-reversion speed, factor exposure); the novel
    contribution is a crowding measure $c_t = W_2(\mu_t, \mu_t^*)$ that quantifies
    distributional disequilibrium. Methods for estimating the unobservable equilibrium
    measure (GMM, Kalman filtering, reduced-form proxies) are provided, and a
    mean-variance portfolio construction with a crowding discount is developed.

    \item \textbf{Empirical validation methodology, demonstrated on synthetic
    S\&P 500-style data (2000--2024)} (Chapter 9). A strategy based on the four signals,
    using a methodology specified for point-in-time 13F holdings and CFTC COT
    data, achieves an out-of-sample Sharpe ratio of 1.91 (net of 5--10bps transaction
    costs) on a synthetic panel constructed with a known embedded predictive signal,
    with a maximum drawdown of $-18.3\%$ and positive skewness (0.32). Statistical
    significance is $p = 0.002$ (block bootstrap, 10{,}000 samples). Performance is robust
    to alternative cost assumptions, calibration windows, discount functions, and
    sub-periods, and incremental over both a na\"ive 13F mean-reversion benchmark
    (Sharpe 0.72) and a 13F residual momentum benchmark (Sharpe 0.89), all evaluated on
    the same synthetic panel. The crowding measure
    $c_t$ is strongly monotonic in subsequent alpha: the top quintile yields average
    monthly returns of 2.8\% versus 0.9\% in the bottom quintile. \textbf{These figures
    demonstrate that the methodology behaves as intended on data with known structure;
    they are not evidence of profitability on real markets}, since the licensed data
    sources (CRSP, Compustat, SEC 13F, CFTC COT) were not available in the environment
    where this thesis was produced. Deployment on real market data is identified as
    future work.
\end{enumerate}

Selected core estimates --- the Gronwall inequality, the Riccati system derivation, the
CFL condition, and the optimality of the linear feedback control --- have been formally
verified in the Lean 4 interactive theorem prover; the code is publicly available. The
thesis concludes with a candid discussion of limitations (regularity assumptions,
subsequential convergence, model-risk dominance in the empirical Sharpe ratio) and a map
of six open problems, including jump-diffusion extensions, online RL with streaming data,
and full formal verification of the pipeline.

All theoretical results hold under the standing assumptions collected in Section 1.9;
these are strong and their practical implications are discussed throughout. The reported
Sharpe ratio is conditional on the LQ model specification and estimation procedures;
model risk dominates statistical risk.

\vspace{1em}
\noindent\textbf{Keywords:} Mean-field games, common noise, Wasserstein distance,
contraction, multi-timescale reinforcement learning, actor-critic, hierarchical
architecture, alpha generation, crowding measure.
\end{abstract}
